Evaluation of the article

### 2024-11-05 pages/paper_evaluation.qmd
docs(paper evaluation): add ispor-sdm reflections, and charts for use of reporting guidelines or previously reported model

```{python}
plot_scatter('STRESS-DES criteria met',
             sho = (11/15)*100,
             hua = (7/16)*100,
             lim = (12/16)*100,
             kim = (12/17)*100,
             ana = (8/15)*100,
             joh = (15/17)*100,
             her = (10/17)*100,
             woo = (13/17)*100)
```

@@

These guidelines have quite a different focus from STRESS-DES. STRESS-DES is very focused on what and how modelling work is performed. These guidelines do cover that, but have quite alot of focus on validity of the work, and good practice (e.g. stating funding, involving stakeholders).

<!--TODO: Add reflections on each of the elements -->

@@

*

@@

*

@@

*

@@

*

@@

*

@@

*

@@

*

@@

*

@@

*

@@

*

@@

*

@@

*

@@

*

@@

*

@@

* Good practice, usually a journal requirement, although some did not have.

@@

*

@@

@@

<!--TODO: Add reflections on this, and relationship between these and then e.g. how many of the reporting guidelines were met -->

```{python}
import plotly.express as px

# For each study, define (a) whether use a previously reported model, and (b) proportion of applicable STRESS-DES fully met
previous_stress =  [
  (0, (17/24)*100),
  (0, (14/23)*100),
  (0, (16/21)*100),
  (0, (15/22)*100),
  (0, (14/21)*100),
  (1, (16/19)*100),
  (0, (18/23)*100),
  (1, (22/24)*100)
]

# Get x and y values
x, y = zip(*previous_stress)

fig = px.scatter(x=x, y=y)
fig.show()
```

@@

<!--TODO: Add reflections on this, and relationship between these and then e.g. how many of the reporting guidelines were met -->

```{python}
import plotly.express as px

# For each study, define (a) whether use a previously reported model, and (b) proportion of applicable STRESS-DES fully met
previous_stress =  [
  (0, (17/24)*100),
  (0, (14/23)*100),
  (0, (16/21)*100),
  (1, (15/22)*100),
  (0, (14/21)*100),
  (1, (16/19)*100),
  (0, (18/23)*100),
  (1, (22/24)*100)
]

# Get x and y values
x, y = zip(*previous_stress)

fig = px.scatter(x=x, y=y)
fig.show()
```

<!--TODO: Improve this plot so (a) it's not a massive gap between the bars, (b) it's a function used for this and above rather than duplicate code, (c) so we have ones for STRESS and ISPOR -->

<!-- TODO: consider, did the previously reported model also use reporting guidelines? -->

>>>>

```{python}
plot_scatter('ISPOR-SDM criteria met',
             sho = (11/15)*100,
             hua = (7/16)*100,
             lim = (12/16)*100,
             kim = (12/17)*100,
             ana = (8/15)*100,
             joh = (15/17)*100,
             her = (10/17)*100,
             woo = (13/17)*100)
```

@@

These guidelines have quite a **different focus** from STRESS-DES. STRESS-DES is very focused on what and how modelling work is performed. These guidelines do cover that, but have quite alot of focus on validity of the work, and good practice (e.g. stating funding, involving stakeholders). This is important, as **different checklists can be important** - and although practice is typically to use one practice, referring to more than one could also be beneficial.

@@

@@

* There was only one study where this was not stated (@anagnostou_facs-charm_2022), though this didn't impact the reproduction itself, for this study I was simply able to run the code and pretty much get the result required with minimal troubleshooting.

@@

* All papers with scenarios described them. As reflected for STRESS-DES, though the description of the scenario can feel clear from the paper, when the code was not provided, it could be quick tricky and time-consuming to work out how to appropriately change the code, in order to implement the scenario.

@@

* Fulfilment didn't impact the reproduction as this is more about interpretation/validity/etc.

@@

* All met, although it's worth noting that I didn't check that *every* single parameter's data source was stated (and indeed, its likely some were not) - simply whether I could identified that at least some were.

@@

* This was quite important for the reproduction as it allows us to check the parameters in the code are correct. When I identified that there were some discrepancies between the article and code, then in cases where a parameter was not given in the article, I couldn't be sure if it was correct or not in the code, as there was nothing to compare against.

@@

* Some don't (but worth noting this doesn't have impact on reproduction - beyond just these being additional values/lines in plot that we are trying to reproduce)

@@

* Some performed, some mentioned but no results presented, and some didn't mention at all. One explained it to not be relevant.

@@

* Rare (only one completed) - but didn't impact reproduction as this is more related to validity

@@

* Rare (only two completed) - but didn't impact reproduction as this is more related to validity

@@

* Rare (only two completed) - although some argue not applicable - but didn't impact reproduction as this is more related to validity

@@

* Generally not applicable - but didn't impact reproduction as this is more related to validity

@@

* More than half did discuss this - but didn't impact reproduction as this is more related to reuse

@@

* Rare (only one mentioned) - but didn't impact reproduction as this is more related to validity

@@

* Good practice, usually a journal requirement, although some did not have - but didn't impact reproduction.

@@

* Most discuss limitations (didn't really assess how comprehensive these were though, except in two cases where they only had a very general or hint towards limitations) - but didn't impact reproduction.

@@

It sometimes took quite a while to find all this information from the articles - and we acknowledge there's a chance that it is provided somewhere in the article but I missed it. For both of these reasons, we see the value in actually attaching a completed reporting checklist, clearly laying out key information about the model and study.

@@

Although this is only a small sample, its interesting to note that the two studies that used reporting guidelines both had the highest proportion of fully met criteria in either reporting guideline.

```{python}
from functions import reporting_scatter

for guide in ['stress', 'ispor']:
    fig = reporting_scatter(
        'Used reporting guidelines', reporting=guide,
        sho=0, hua=0, lim=0, kim=0, ana=0, joh=1, her=0, woo=1)
    fig.show()
```

@@

Again, a small sample, but this time a weaker pattern. We note that the two studies that used reporting guidelines are the same that are previously reported models here, alongside one other study which was previously reported but did not use reporting guidelines in this instance, and has a lower proportion of criteria that were fully met.

```{python}
for guide in ['stress', 'ispor']:
    fig = reporting_scatter(
        'Model reported previously', reporting=guide,
        sho=0, hua=0, lim=0, kim=1, ana=0, joh=1, her=0, woo=1)
    fig.show()
```
